Reanalysis Step 4: Outcome Analysis with Three-Way Interaction

### 2025-08-19 reanalysis-2-multiple_imputation.qmd
bug fixes on mi

df_mice <- df_for_imputation |>
  dplyr::select(dplyr::all_of(imputation_vars)) |>
  mutate(
    # Convert person_dr to ordered factor for polr imputation method
    person_dr = factor(person_dr, levels = c(0, 1, 2), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Lock factor levels for categorical variables to ensure consistency
    gender_cat = factor(gender_cat),
    race_ethnic_cat = factor(race_ethnic_cat),
    insurance_cat = factor(insurance_cat)
  )
```

@@

# Don't impute these variables
# Method = "" means: don't impute (but still use as predictor EXCEPT for ID)
custom_methods["e_mrn_deidentified"] <- ""  # ID - never impute, never use as predictor
custom_methods["ever_lapse_binary"] <- ""  # Complete - use as predictor only
custom_methods["glaucoma_bef_hitplus_cat"] <- ""  # Already recoded to 0/1
custom_methods["otherretina_bef_hitplus_cat"] <- ""  # Already recoded to 0/1
custom_methods["gender_cat"] <- ""  # Likely complete
custom_methods["other_inject"] <- ""  # Likely complete
custom_methods["anti_VEGF"] <- ""  # Likely complete  
custom_methods["focal_laser_flag"] <- ""  # Likely complete
custom_methods["PRP_flag"] <- ""  # Likely complete

@@

# Don't use ID as predictor AND don't impute it
pred_matrix[, "e_mrn_deidentified"] <- 0
pred_matrix["e_mrn_deidentified", ] <- 0

@@

# Add back any variables not included in imputation or transformation
# Get the full original dataset
df_original_full <- df_for_imputation

# Variables to add back (those not in imputation model or created by transformations)
vars_in_imputed <- names(imputed_datasets[[1]])
vars_to_add <- setdiff(names(df_original_full), vars_in_imputed)
```

@@

```{r add-back-vars}
if(length(vars_to_add) > 0) {
  # Add these variables back to each imputed dataset
  imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
    imp_data <- imp_data |>
      left_join(
        df_original_full |> 
          dplyr::select(e_mrn_deidentified, dplyr::all_of(vars_to_add)),
        by = "e_mrn_deidentified",
        relationship = "many-to-many"  # Expected: multiple eyes per patient
      )

>>>>

df_mice <- df_for_imputation |>
  mutate(
    # Create a unique row identifier to preserve exact row matching
    row_id = row_number()
  ) |>
  dplyr::select(row_id, dplyr::all_of(imputation_vars)) |>
  mutate(
    # Convert person_dr to ordered factor for polr imputation method
    person_dr = factor(person_dr, levels = c(0, 1, 2), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Lock factor levels for categorical variables to ensure consistency
    gender_cat = factor(gender_cat),
    race_ethnic_cat = factor(race_ethnic_cat),
    insurance_cat = factor(insurance_cat)
  )
```

@@

# Don't impute these variables
# Method = "" means: don't impute (but still use as predictor EXCEPT for ID)
custom_methods["row_id"] <- ""  # Row ID - never impute, never use as predictor
custom_methods["e_mrn_deidentified"] <- ""  # Patient ID - never impute, never use as predictor
custom_methods["ever_lapse_binary"] <- ""  # Complete - use as predictor only
custom_methods["glaucoma_bef_hitplus_cat"] <- ""  # Already recoded to 0/1
custom_methods["otherretina_bef_hitplus_cat"] <- ""  # Already recoded to 0/1
custom_methods["gender_cat"] <- ""  # Likely complete
custom_methods["other_inject"] <- ""  # Likely complete
custom_methods["anti_VEGF"] <- ""  # Likely complete  
custom_methods["focal_laser_flag"] <- ""  # Likely complete
custom_methods["PRP_flag"] <- ""  # Likely complete

@@

# Don't use IDs as predictors AND don't impute them
pred_matrix[, "e_mrn_deidentified"] <- 0
pred_matrix["e_mrn_deidentified", ] <- 0
pred_matrix[, "row_id"] <- 0
pred_matrix["row_id", ] <- 0

@@

# Add back any variables not included in imputation or transformation
# Get the full original dataset with row IDs
df_original_full <- df_for_imputation |>
  mutate(row_id = row_number())

# Variables to add back (those not in imputation model or created by transformations)
vars_in_imputed <- names(imputed_datasets[[1]])
# Remove row_id from comparison since it's our join key
vars_to_add <- setdiff(names(df_original_full), c(vars_in_imputed, "row_id"))
```

@@

```{r add-back-vars}
if(length(vars_to_add) > 0) {
  # Add these variables back to each imputed dataset using row_id for exact matching
  imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
    # Store original row count
    orig_rows <- nrow(imp_data)

    # Join using row_id for exact 1:1 matching
    imp_data <- imp_data |>
      left_join(
        df_original_full |> 
          dplyr::select(row_id, dplyr::all_of(vars_to_add)),
        by = "row_id"
      )

    # Verify no row duplication
    if (nrow(imp_data) != orig_rows) {
      stop("Join changed row count from ", orig_rows, " to ", nrow(imp_data), 
           ". This should not happen with row_id join.")
    }

    # Remove row_id as it's no longer needed
    imp_data <- imp_data |> select(-row_id)

### 2025-08-19 reanalysis-4-outcome_analysis.qmd
bug fixes on mi

```{r pool-results}
# Use mitools for robust pooling implementation
# MIcombine properly implements Rubin's rules with Barnard-Rubin adjustment
pooled_model <- MIcombine(three_way_models)

>>>>

```{r pool-results}
# Use mitools for robust pooling implementation
# MIcombine properly implements Rubin's rules with Barnard-Rubin adjustment
pooled_model <- MIcombine(models)
